\documentclass[12pt]{article}

\usepackage[thinlines]{easytable}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{fullpage}

\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Requests}{\mathfrak{R}}
\newcommand{\T}{\mathcal{T}}
\newcommand{\I}{\mathcal{I}}
\newcommand{\E}{\mathcal{E}}

\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\Exp}{\mathbf{E}}

\DeclareMathOperator*{\maximize}{maximize}
\DeclareMathOperator*{\minimize}{minimize}
\DeclareMathOperator*{\subjectto}{subject\ to}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}

\begin{document}

\title{Pacing}
\author{D\'avid P\'al}
\date{June 20, 2026}

\maketitle

\begin{abstract}
We show convergence results for the pacing problem.
\end{abstract}

\section{Introduction}

Consider a sequence of requests $R_1, R_2, \dots, R_T$. For each request, we
have to choose one of $N$ possible actions. Each request is an $N \times (M+1)$
matrix with entries in the interval $[-1,1]$.\footnote{As long as the entries
are bounded, they can be linearly rescaled to the interval $[-1,1]$.} In other
words, we assume that requests are elements of the set $\Requests = [-1,1]^{N
\times (M+1)}$.  We make a simple assumption that the requests are sampled
i.i.d. from a fixed, but unknown, distribution $D$.\footnote{The probability
distribution is defined on the Borel $\sigma$-algebra generated by the open sets
of $\Requests$.} We would like to use a (possibly randomized) policy $\pi$ to
choose the action for each request. A randomized policy $\pi$ is a measurable
function $\pi:\Requests \to \Delta_N$, where $\Delta_N$ is the probability
simplex defined as
$$
\Delta_N = \left\{ x \in \R^N ~:~ x_1, x_2, \dots, x_N \ge 0, \sum_{i=1}^N x_i = 1  \right\} \: .
$$
Measurability of $\pi$ is defined with respect to the $\sigma$-algebras defined
by the Borel sets of $\Requests$ and $\Delta_N$ respectively. We denote by $\Pi$
the set of all policies.

We would like to come up with a policy that is ``close'' to an optimal solution
of the optimization problem
\begin{equation}
\label{equation:sample-policy-optimization}
\begin{aligned}
\maximize_{\pi \in \Pi} \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot \pi(R_t)_i \\
\subjectto \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi(R_t)_i \le B_j & \text{for all $j=1,2,\dots,M$}
\end{aligned}
\end{equation}

The value $R_{t,i,0}$ is the reward collected for $t$-th request if we make
decision $i$ for it. The value $R_{t,i,j}$ for $j = 1, 2, \dots, M$ is the
$j$-th business metric for the $t$-th request if we make decision $i$ for it.
The right-hand sides of the constraints $B_1, B_2, \dots, B_M$ are given to us
ahead of time.

For example, at Uber, a request can be a request by a customer to get a ride and
Uber needs to choose one of $N$ possible prices for the request. $R_{t,i,0}$
could be the probability of the trip happening, $R_{t,i,1}$ could be the price
paid by the rider, $R_{t,i,2}$ could be the price paid to the driver, etc. The
right-hand sides of the constraints are the business objectives Uber wishes to
achieve over the period of say, one day or one week.

In practical applications, the number of requests, $T$, is not known ahead of time.
Instead, it is forecasted by a statistical or a machine learning model. We
ignore the issue of forecasting $T$ and we assume that $T$ is known precisely
ahead of time.

In practical applications, the entries of the matrix $R_t$ are often predicted
by machine learning models just before the action needs to be chosen. These
predictions are only proxies for the actual reward and business metrics. The
actual reward and business metrics are revealed only after the decision has been
made, sometimes with a significant delay. However, in order to keep our model
simple, we ignore the difference between the predicted and actual reward and
business metrics.

In practice, the requests change over time. In other words, the sequence of
requests has various seasonal patterns, often at multiple time scales.
Nevertheless, the i.i.d. assumption is a good approximation for short enough
time periods (e.g. one hour, one day, or one week).

The optimization problem~\eqref{equation:sample-policy-optimization} cannot be
solved directly, since we don't have all the coefficients $R_{t,i,j}$ available
upfront. Instead, the requests arrive one at a time, and for each request we
need to choose a probability vector in the probability simplex $\Delta_N$. The
probability vector for each request is computed by applying a pre-computed
policy $\pi$ to the request. A good pre-computed policy $\pi$ can be found by
finding the optimal solution to the following optimization problem:
\begin{equation}
\label{equation:distribution-policy-optimization}
\begin{aligned}
\maximize_{\pi \in \Pi} \qquad & \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,0} \right] \\
\subjectto \qquad & \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,j} \right] \le B_j & \text{for all $j=1,2,\dots,M$} \\
\end{aligned}
\end{equation}

However, the optimization problem~\eqref{equation:sample-policy-optimization}
cannot be solved directly either, since the distribution $D$ is unknown.
However, we assume that we have a historical sample of requests available that
we can use. We assume that the historical sample is generated i.i.d. from $D$ as
well. Using the historical sample, we formulate an optimization problem. The
problem is identical to \eqref{equation:sample-policy-optimization} with the
only difference being that $R_1, R_2, \dots, R_T$ refers to the historical
sample and the sample size $T$ refers to the size of the historical sample.

We study the relationship between the optimization problems
\eqref{equation:sample-policy-optimization} and
\eqref{equation:distribution-policy-optimization}. We show that good solutions
to both problems lie in a much smaller class of policies, which we call
Lagrangian policies. We show that a Lagrangian policy that is a good solution
for one of the optimization problems is a good solution for the other.
Lagrangian policies naturally arise from the formulation of the optimization
problem~\eqref{equation:sample-policy-optimization} as a linear program. The
dual of the linear program plays a central role in our analysis.

We focus on a regime where the number of requests $T$ is much greater than both
$N$ and $M$. The assumption that the entries of a request matrix $R_t$ lie in
$[-1,1]$ translates to an assumption that each request has a negligible impact
on the total (or average) business metrics.

We use the following notation for the rest of the paper. For $i=1,2,\dots,N$, we
define $e_i$ to be the vector in $\R^N$ with $i$-th entry equal to $1$ and all
other entries equal to zero, that is,
$$
e_i = \underbrace{(0,\dots,0,1,0,\dots,0)}_{\text{$1$ is at position $i$}} \: .
$$
We denote by $\|\cdot\|_2$ the Euclidean norm on $\R^{M+1}$. That is, if $\lambda = (\lambda_0, \lambda_1, \dots, \lambda_M) \in \R^{M+1}$, then
$$
\|\lambda\|_2 = \sqrt{\lambda_0^2 + \lambda_1^2 + \lambda_2^2 + \dots + \lambda_M^2} \: .
$$

\section{Lagrangian policies}

We consider policies parameterized by $M+1$ real numbers $\lambda_0, \lambda_1,
\dots, \lambda_M$. As will become clear later, the $\lambda_1, \dots, \lambda_M$
can be interpreted as the dual variables associated with the constraints of the
optimization problems. However, for now, they are arbitrary real numbers. For
any $(M+1)$-dimnesional vector $\lambda = (\lambda_0, \lambda_1, \dots,
\lambda_M) \in \R^{M+1}$ we define a policy $\pi_\lambda:\Requests \to \Delta_N$
as follows. For any request $R \in \Requests$, we define
\begin{equation}
\label{equation:lagrangian-policy}
\pi_\lambda(R) = e_{i^*}
\qquad \text{where} \qquad
i^* = \argmax_{i=1,2,\dots,N} \lambda_0 R_{i,0} + \lambda_1 R_{i,1} + \lambda_2 R_{i,2} + \dots + \lambda_M R_{i,M} \: .
\end{equation}
The ties in the $\argmax$ are broken towards the smaller index $i$. We call a
policy of this form a \emph{Lagrangian policy}. The reason for this name will
become clear once we formulate the optimization problem as a linear program and
consider its dual.

We make three remarks about the Lagrangian policies. First, observe that
Lagrangian policies are both deterministic and measurable.

Second, note that that the policy corresponding to the all-zero vector $\lambda
= (0, 0, \dots, 0)$ chooses the first decision for any request. This policy is
uniteresting. For technical reasons and/or notational convenience, we include it
in certain theorems and exclude in others. However, as it will turn there exists
a Lagrangian policy $\pi_\lambda$ with $\lambda \neq (0, 0, \dots, 0)$ that is
close to an optimal policy the optimization problems
\eqref{equation:sample-policy-optimization} and
\eqref{equation:distribution-policy-optimization}.

Third, note that for any positive real number $c > 0$ and any vector $\lambda
\in \R^{M+1}$, the policies $\pi_{c \lambda}$ and $\pi_\lambda$ are identical.
Thus, without loss of generality, we can assume that $\|\lambda\|_2 \le 1$. If
$\lambda = 0$ is excluded, we can assume that $\|\lambda\|_2 = 1$. Depending on
the context, the Lagrangian policies are parameterized by $\lambda$ lying in
either $\R^{M+1}$, $\R^{M+1} \setminus \{0\}$, unit ball, or unit sphere. The
advantage of the unit ball and the unit sphere is that they are compact sets.

\section{Failure of deterministic policies}

In general, determistic policies are not good solutions for certain sequences of
requests. In this section, we provide the simplest possible example that
illustrates the point. Suppose $N=2$ and $M=1$ and $T$ can be atrbitrary. For
all requests,
$$
R_t
= \begin{pmatrix} R_{t,1,0} & R_{t,1,1} \\ R_{t,2,0} & R_{t,2,1} \end{pmatrix}
= \begin{pmatrix} 1 & 1 \\ 0 & 0 \end{pmatrix}
$$
and $B_1 = \frac{1}{2}$. Any optimal policy $\pi^*$ for this sequence of
requests satisfies
\begin{equation}
\label{equation:uniform-random-policy-on-two-outcomes}
\pi^*(R_t) = \left( \frac{1}{2}, \frac{1}{2} \right) \: .
\end{equation}
The output of a policy for requests other than $R_t$ does not play any role and
it can be arbitrary. More importantly, any policy that does not satisfy
condition \eqref{equation:uniform-random-policy-on-two-outcomes} is suboptimal.

However, as we will see later, Lagrangian policies are almost optimal assuming
that the requests are sampled i.i.d. from a distribution with a probability
density.

\section{Distributions with a density}

So far we avoided to make any assumption on the distribution $D$. For technical
reasons, it is convenient to assume that $D$ has a probability density with
respect to the Lebesgue measure on $\Requests = [-1,1]^{N \times (M+1)}$. The
fundamental technical reason for the existence of a probability density is to
avoid having ties in the argmax in equation \eqref{equation:lagrangian-policy}
defining Lagrangian policies.

If the distribution $D$ does not have a probability density, we can always
create a ``nearby'' distribution $D'$ that does. One way of doing it is by
adding a small independent noise to each entry $R_{t,i,j}$ of the requests.
Namely, we replace $R_{t,i,j}$ with
$$
R'_{t,i,j} = (1-\epsilon)R_{t,i,j} + \epsilon Z_{t,i,j}
\qquad \text{for $t=1,2,\dots,T$, $i=1,2,\dots,N$, $j=0,1,2,\dots,M$}
$$
where $\epsilon \in (0,1)$ is a small positive number and the variables
$Z_{t,i,j}$ drawn i.i.d. from the uniform distribution over $[-1,1]$ and are
independent of the random variables $R_1, R_2, \dots, R_T$. The value of
$\epsilon$ can be arbitrily small. The value of the order $\epsilon =
\Theta(1/\sqrt{T})$ is commensurate with other errors that appear in our
analysis.

\section{Uniform convergence for Lagrangian policies}

The first major result that we prove is that for all Lagrangian policies the
objective functions and the corresponding constraints of the two optimization
problems \eqref{equation:sample-policy-optimization} and
\eqref{equation:distribution-policy-optimization} converge to one another as the
sample size $T$ increases. This result holds for any distribution $D$ with or
without a probability density. The strength of the result is that the convergence
is uniform over all Lagrangian policies. The proof is a standard application
of statistical learning theory; see~\cite{Devroye-Lugosi-2001}.

\begin{theorem}[Uniform convergence]
\label{theorem:uniform-convergence}
Let $D$ be any distribution over $\Requests$. Let $\delta \in (0,1)$ and $T$ be
a positive integer. Suppose $R, R_1, R_2, \dots, R_T$ is an i.i.d. sample from
$D$. Then, with probability at least $1-\delta$, for all $\lambda \in \R^{M+1}$ and for all $j=0,1,2,\dots,M$,
\begin{multline*}
\left| \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi_\lambda(R)_i \cdot R_{i,j}  \right]  -  \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i  \right|
\\
\le \sqrt{\frac{8(M+1) (\ln(4T) + 2 \ln N)}{T}} + \sqrt{ \frac{2\log(1/\delta)}{T} } \: .
\end{multline*}
\end{theorem}

\begin{proof}
Let $R, R_1, R_2, \dots, R_T$ be an i.i.d. sample from $D$ of size $T+1$.
Consider the random variable
$$
Z = \sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}}
\left| \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi_\lambda(R)_i \cdot R_{i,j}  \right] - \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i \right| \: .
$$
The random variable $Z$ is a function of $R_1, R_2, \dots, R_T$. It satisfies
the bounded difference property. Namely, if a single $R_t$ changes, then $Z$
changes by at most $2/T$. McDiarmid's inequality implies that for any $\epsilon > 0$,
$$
\Pr\left[ Z \ge \Exp[Z] + \epsilon \right] \le e^{-T\epsilon^2/2} \: .
$$
We set $\delta = e^{-T\epsilon^2/2}$ and solve for $\epsilon$. We get that with
probability at least $1-\delta$,
$$
Z \le \Exp[Z] + \sqrt{ \frac{2\log(1/\delta)}{T} } \: .
$$
It remains to upper bound $\Exp[Z]$. We use the symmetrization trick. Let $R'_1,
R'_2, \dots, R'_T$ be an i.i.d. sample from $D$ of size $T$. Let $\sigma_1,
\sigma_2, \dots, \sigma_T$ be i.i.d. Rademacher random variables, i.e.,
$\Pr[\sigma_t = 1] = \Pr[\sigma_t = -1] = 1/2$ for all $t=1,2,\dots,T$. We
assume that $\sigma_1, \sigma_2, \dots, \sigma_T$, $R$, $R_1, R_2, \dots, R_T$,
$R'_1, R'_2, \dots, R'_T$ are mutually independent. We have
\begin{align*}
\Exp[Z] & = \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \Exp \left[ \sum_{i=1}^N \pi_\lambda(R)_i \cdot R_{i,j}  \right] - \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i \right| \right] \\
& = \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \Exp \left[ \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N \pi_\lambda(R'_t)_i \cdot R'_{t,i,j}  \right] - \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i \right| \right] \\
& \le \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N \pi_\lambda(R'_t)_i \cdot R'_{t,i,j} - \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i \right| \right] \\
& = \frac{1}{T} \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \left(\pi_\lambda(R'_t)_i \cdot R'_{t,i,j} - R_{t,i,j} \cdot \pi_\lambda(R_t)_i \right) \right| \right] \\
& \le \frac{1}{T} \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \cdot \pi_\lambda(R'_t)_i \cdot R'_{t,i,j} \right| + \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \cdot \pi_\lambda(R_t)_i \cdot R_{t,i,j} \right|  \right] \\
& \le \frac{2}{T} \Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \cdot \pi_\lambda(R_t)_i \cdot R_{t,i,j}  \right| \right]
\end{align*}

Condition on the event $R_1=r_1, R_2=r_2, \dots, R_T=r_T$ and consider the set of $T$-dimensional vectors
$$
V = \left\{ \left(\sum_{i=1}^N \pi_\lambda(r_1)_i \cdot r_{1,i,j}, \dots, \sum_{i=1}^N \pi_\lambda(r_T)_i \cdot r_{T,i,j} \right) ~:~ \lambda \in \R^{M+1}, j=0,1,2,\dots,M \right\} \: .
$$
The set $V$ is finite. We reduce the supremum over an uncountably infinite set to a maximum over the finite set $V$,
$$
\Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}} \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \cdot \pi_\lambda(R_t)_i \cdot R_{t,i,j}  \right| ~\middle|~ R_1=r_1, R_2=r_2, \dots, R_T=r_T \right]
= \Exp\left[\max_{v \in V} \left| \sum_{t=1}^T \sigma_t v_t \right| \right]  \: .
$$
Consider any vector $v = (v_1, v_2, \dots, v_T)$ in $V$. Any entry $v_t$ of $v$ satisfies
$$
|v_t| = \left|\sum_{i=1}^N \pi(r_t)_i \cdot r_{t,i,j}\right| \le \left(\sum_{i=1}^N \pi(r_t)_i\right) \cdot  \max_{i=1,2,\dots,N} |r_{t,i,j}| \le 1 \: .
$$
This implies that the random variable $Y_v = \sum_{t=1}^T \sigma_t v_t$ is a
$\sigma$-sub-Gaussian random variable with $\sigma=\sqrt{T}$. That is, $Y_v$
satisfies
$$
\Exp[e^{sY_v}] \le e^{s^2T/2} \qquad \text{for all $s \in \R$.}
$$
Let $K$ be such that $|V| \le K$ for any choice $r_1, r_2, \dots, r_T \in
\Requests$. Lemma~\ref{lemma:maximum-of-sub-gaussian-variables}
implies that
$$
\Exp\left[\max_{v \in V} \left| \sum_{t=1}^T \sigma_t v_t \right| \right] \le \sqrt{2T \log(2K)}
$$
Thus, taking expectations over $R_1, R_2, \dots, R_T$, we get that
$$
\Exp\left[\sup_{\substack{\lambda \in \R^{M+1} \\ j=0,1,2,\dots,M}}  \left| \sum_{t=1}^T \sum_{i=1}^N \sigma_t \cdot \pi_\lambda(R_t)_i \cdot R_{t,i,j}  \right|  \right] \le \sqrt{2T \log(2K)} \: .
$$
Therefore,
$$
\Exp[Z] \le \sqrt{\frac{8 \log(2K)}{T}} \: .
$$

It remains to find an upper bound $K$ on the cardinality of the set $V$. Let
$$
U = \left\{ \left(\pi_\lambda(r_1), \dots, \pi_\lambda(r_T) \right) ~:~ \lambda \in \R^{M+1} \right\} \: .
$$
Clearly, $|V| \le (M+1) |U|$. It remains to upper bound the cardinality of $U$.

Consider any $\lambda \in \R^{M+1}$ and the corresponding Lagrangian policy
$\pi_\lambda$. The output of the policy $\pi_\lambda$ on a request $r_t$ is
determined by $\binom{N}{2}$ indicators
$$
\mathbf{1} \left[\lambda_0 r_{t,i,0} - \lambda_1 r_{t,i,1} - \dots - \lambda_M r_{t,i,M} \ge \lambda_0 r_{t,i',0} - \lambda_1 r_{t,i',1} - \dots - \lambda_M r_{t,i',M}\right]
$$
indexed by pairs $i,i'$ such that $1 \le i < i' \le N$. Using vector notation, we can write the indicator as
$$
\mathbf{1} \left[ \lambda^T (r_{t,i} - r_{t,i'}) \ge 0 \right] \: .
$$
Let
$$
\kappa_\lambda(r_t) = (\mathbf{1} \left[ \lambda^T (r_{t,1} - r_{t,2}) \ge 0 \right], \mathbf{1} \left[ \lambda^T (r_{t,1} - r_{t,3}) \ge 0 \right], \dots, \mathbf{1} \left[ \lambda^T (r_{t,N-1} - r_{t,N}) \ge 0 \right])
$$
be the vector of these $\binom{N}{2}$ indicators. Let
$$
W = \left\{ (\kappa_\lambda(r_1), \kappa_\lambda(r_2), \dots, \kappa_\lambda(r_T)) ~:~ \lambda \in \R^{M+1} \right\} \: .
$$
Clearly, $|U| \le |W|$. It remains to upper bound the cardinality of $W$. The
cardinality of $W$ is upper bounded by the number of behaviors of homogeneous
hyperplanes in $\R^{M+1}$ on $\binom{N}{2} T$ points. The Vapnik-Chervonenkis
dimension of homogeneous hyperplanes in $\R^{M+1}$ is $M+1$.\footnote{The
Vapnik-Chervonenkis dimension is not affacted by the inclusion of the constant
$1$ predictor that corresponds to $\lambda = 0$.} Thus, the number of behaviors
of homogeneous hyperplanes in $\R^{M+1}$ on $\binom{N}{2} T$ points and thus the
cardinality of $W$ is at most
$$
|W| \le \sum_{k=0}^{M+1} \binom{T \binom{N}{2}}{k} \le \left(T \binom{N}{2} + 1 \right)^{M+1} \le (N^2T)^{M+1} \: .
$$
Therefore, the cardinality of $V$ is upper bounded as
$$
|V| \le (M+1) |W| \le (M+1) \left(N^2T \right)^{M+1} \le 2^{M+1} \left(N^2T \right)^{M+1} = \left(2N^2T \right)^{M+1} \: .
$$
Using the value $K=\left(2N^2T \right)^{M+1}$, we get that
$$
\Exp[Z] \le \sqrt{\frac{8 \ln(2K)}{T}} \le \sqrt{\frac{8 \ln((4N^2T)^{M+1})}{T}} = \sqrt{\frac{8(M+1) (\ln(4T) + 2 \ln N)}{T}} \: .
$$
\end{proof}


\section{Linear programs}

The optimization problem \eqref{equation:sample-policy-optimization} is
equivalent to the linear program below provided that the requests $R_1, R_2,
\dots, R_T$ are pairwise distinct. If the requests are sampled i.i.d. from a
probability with a density, they are pairwise distinct with probability one.
\begin{equation}
\label{equation:primal-linear-program}
\begin{aligned}
\maximize_{x_{1,1}, \dots, x_{T,N} \in \R} \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot x_{t,i} \\
\subjectto \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot x_{t,i} \le B_j && \qquad \text{for all $j=1,2,\dots,M$} \\
& \sum_{i=1}^N x_{t,i} = 1 && \qquad \text{for all $t=1,2,\dots,T$} \\
& x_{t,i} \ge 0 && \qquad \text{for all $i=1,2,\dots,N$} \\
& && \qquad \text{and all $t=1,2,\dots,T$}
\end{aligned}
\end{equation}
The vector $(x_{t,1}, x_{t,2}, \dots, x_{t,N})$ encodes the policy output
$\pi(R_t)$. The output of the policy on inputs other than $R_1, R_2, \dots, R_T$
is unspecified. We refer to the linear program
\eqref{equation:primal-linear-program} as \emph{the primal linear program}.

The dual of the primal linear program \eqref{equation:primal-linear-program}
\begin{equation}
\label{equation:dual-linear-program}
\begin{aligned}
\minimize_{\substack{\mu_1, \dots, \mu_T \in \R \\ \lambda_1, \dots, \lambda_M \in \R}} \qquad & \sum_{t=1}^T \mu_t + \sum_{j=1}^M B_j \lambda_j \\
\subjectto \qquad
& \mu_t + \frac{1}{T} \sum_{j=1}^M R_{t,i,j} \lambda_j  \ge \frac{1}{T} R_{t,i,0} && \qquad \text{for all $i=1,2,\dots,N$} \\
& && \qquad \text{and all $t=1,2,\dots,T$} \\
& \lambda_j \ge 0 && \qquad \text{for all $j=1,2,\dots,M$}
\end{aligned}
\end{equation}
We refer to this linear program as \emph{the dual linear program}.

We prove several results about these linear programs. We start with easy
observations first. The feasible region of the primal linear program is bounded.
Indeed, the feasible region is a subset of $\left(\Delta_N\right)^T \subset
\R^{N \times T}$. The primal program is either feasible and has an optimal
solution, or it is infeasible; it cannot be unbounded. By the strong duality
theorem, if the primal program is feasible, the dual program is feasible as well
and has an optimal solution. However, even if the primal program is feasible,
the dual program can be have an unbounded feasible set, and even unbounded set
of optimal solutions.

Next, we prove that the optimal values of the dual variables $\mu_1, \dots,
\mu_T$ and the optimal values of ``some'' primal variables, $x_{1,1}, \dots,
x_{T,N}$, can be computed from the optimal values of only $\lambda_1, \dots,
\lambda_M$. We then prove that, in fact, the optimal values of ``almost all''
primal variables $x_{1,1}, \dots, x_{T,N}$ can be recovered from the optimal
values of $\lambda_1, \dots, \lambda_M$ provided that the requests are in a
``general position'', a term that we define precisely later. We show that if the
requests are sampled i.i.d. from a probability distribution with a probability
density then they are in general position with probability one. In summary, the
optimal solutions of both linear programs can be essentially encoded using only
$M$ parameters $\lambda_1, \dots, \lambda_M$. Finally, we show that the
resulting Lagrangian policy $\pi_\lambda$ is close to the optimal solution of
the primal program.

\begin{proposition}[Computing $\mu$]
\label{proposition:computing_mu}
Let $\mu_1, \dots, \mu_T$, $\lambda_1, \dots, \lambda_M$ be any optimal solution
of the dual program. Then,
$$
\mu_t = \frac{1}{T} \max_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right) \: .
$$
\end{proposition}

\begin{proof}
The dual constraints imply that
$$
\mu_t \ge \frac{1}{T} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right) \: .
$$
Taking maximum over $i$, we get
$$
\mu_t \ge \frac{1}{T} \max_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right) \: .
$$
Any value of $\mu_t$ that satisfies this inequality is a feasible solution of
the dual program. Since $\mu_t$ appears with coefficients $1$ in the dual
objective function, the inequality must hold with equality.
\end{proof}

\begin{proposition}[Computing primal from dual]
\label{proposition:computing_primal_from_dual}
Let $\mu_1, \dots, \mu_T$, $\lambda_1, \dots, \lambda_M$ be any optimal solution
of the dual program. Let $x_{1,1}, \dots, x_{t,N}$ be any optimal solution of
the primal program. For any index $t$, $1 \le t \le T$, if the argmax
$$
i(t) = \argmax_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right)
$$
is uniquely defined, then $x_{t,i(t)} = 1$ and $x_{t,i} = 0$ for all $i \neq i(t)$.
\end{proposition}

\begin{proof}
Suppose
$$
i(t) = \argmax_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right)
$$
is uniquely defined. Proposition~\ref{proposition:computing_mu} implies that
$$
\mu_t = \frac{1}{T} \max_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right) = \frac{1}{T} \left( R_{t,i(t),0} - \sum_{j=1}^M R_{t,i(t),j} \lambda_j \right) \: .
$$
These two facts together imply that
$$
\mu_t > \frac{1}{T} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right) \qquad \text{for all $i \neq i(t)$.}
$$
Equivalently,
$$
\mu_t + \frac{1}{T} \sum_{j=1}^M R_{t,i,j} \lambda_j > \frac{1}{T} R_{t,i,0} \qquad \text{for all $i \neq i(t)$.}
$$
By complementary slackness,
$$
x_{t,i} = 0 \qquad \text{for all $i \neq i(t)$}.
$$
The primal constraint $\sum_{i=1}^N x_{t,i} = 1$ then implies that $x_{t,i(t)} = 1$.
\end{proof}

\begin{definition}[General position]
\label{definition:general-position}
Let $R_1, R_2, \dots, R_T \in \Requests$ be requests. The requests are in a
\emph{general position} if for every choice $(i_1,i_1'), (i_2,i_2'), \dots,
(i_T,i_T')$ of pairs of integer indices such that $1 \le i_1 < i_1' \le N$, $1
\le i_2 < i_2' \le N$, \dots, $1 \le i_T < i_T' \le N$, from the $T$ vectors
$R_{1,i_1} - R_{1,i_1'}$, $R_{2,i_2} - R_{2,i_2'}$, $R_{T,i_2} - R_{T,i_T'}$ in
$\R^{M+1}$ any $M+1$ of them are linearly independent.
\end{definition}

\begin{theorem}[General position]
\label{theorem:general-position}
Suppose that the requests $R_1, R_2, \dots, R_T$ are sampled i.i.d. from a
probability distribution $D$ with a probability density. Then, the requests are
in general position with probability one.
\end{theorem}

\begin{proof}
For any choice of indices $\I = ((i_1,i_1'), (i_2,i_2'), \dots, (i_T,i_T'))$
such that $1 \le i_1 < i_1' \le N$, $1 \le i_2 < i_2' \le N$, \dots, $1 \le i_T
< i_T' \le N$, and any choice of indices $\T = (t_1, t_2, \dots, t_{M+1})$ such
that $1 \le t_1 < t_2 < \dots < t_{M+1} \le T$, consider the $(M+1) \times
(M+1)$ matrix
$$
A_{\I,\T} = \begin{pmatrix}
R_{t_1,i_{t_1},0} - R_{t_1,i_{t_1}',0} &  \dots & R_{t_1,i_{t_1},M} - R_{t_1,i_{t_1}',M} \\
R_{t_2,i_{t_2},0} - R_{t_2,i_{t_2}',0} &  \dots & R_{t_2,i_{t_2},M} - R_{t_2,i_{t_2}',M} \\
 \vdots & & \vdots \\
R_{t_{M+1},i_{t_{M+1}},0} - R_{t_{M+1},i_{t_{M+1}}',0} & \dots & R_{t_{M+1},i_{t_{M+1}},M} - R_{t_{M+1},i_{t_{M+1}}',M}
\end{pmatrix}
$$
The matrix $A_{\I,\T}$ is sampled from a distribution that has a density with respect to
the Lebesgue measure on $\R^{(M+1) \times (M+1)}$. Since the Lebesgue measure of
singular matrices is zero, the probability that $A_{\I,\T}$ is singular is zero as well.

The event that the requests are \emph{not} in a general position is the union of
the events that the matrix $A_{\I,\T}$ is singular over all possible choices of
$\I$ and $\T$. Since there are only finitely of these choices for $\I$ and $\T$,
the probability of the union is zero.
\end{proof}

\begin{lemma}[Overdetermined system of linear equations]
\label{lemma:overdetermined-system-of-linear-equations}
Suppose that $A \in \R^{L \times M}$ and $b \in \R^L$. If the augmented
$L \times (M+1)$ matrix $[A | b]$ has rank $M+1$, then the system of linear
equations $A \lambda = b$ has no solution.
\end{lemma}

\begin{proof}
Suppose by contradiction that the system of linear equations $A \lambda = b$ has
a solution. Then, there exists $\lambda \in \R^M$ such that $A \lambda = b$. In
other words, $b$ can be expressed as a linear combination of the columns of $A$.
Thus, the columns of the augmented matrix $[A | b]$ are linearly dependent. But
then $[A | b]$ has rank at most $M$. This contradicts the assumption that the
rank is $M+1$.
\end{proof}

\begin{theorem}[Few ties]
\label{theorem:few-ties}
Suppose that the requests $R_1, R_2, \dots, R_T$ are in general position. Let
$\lambda_1, \dots, \lambda_M$ be any real numbers. With probability one, for at
least $T-M$ values of the index $t$, the argmax
\begin{equation}
\label{equation:argmax}
\argmax_{i=1,2,\dots,N} \left( R_{t,i,0} - \sum_{j=1}^M R_{t,i,j} \lambda_j \right)
\end{equation}
is uniquely defined.
\end{theorem}

\begin{proof}
Let $\T$ be the set of indices $t$ for which the argmax is not unique. Formally,
$$
\T = \left\{ t \in \Z ~:~ 1 \le t \le T, \text{argmax \eqref{equation:argmax} is \emph{not} unique} \right\}
$$
By the definition of $\T$, for every $t \in \T$
there exists a pair of indices $(i_t,i'_t)$, $i_t < i'_t$ such that
$$
R_{t,i_t,0} - \sum_{j=1}^M R_{t,i_t,j} \lambda_j = R_{t,i'_t,0} - \sum_{j=1}^M R_{t,i'_t,j} \lambda_j \: .
$$
In other words, $\lambda_1, \lambda_2, \dots, \lambda_M$ is a solution of the
system of $|\T|$ linear equations,
$$
\sum_{j=1}^M (R_{t,i_t,j} - R_{t,i'_t,j}) \lambda_j = R_{t,i_t,0} - R_{t,i'_t,0} \qquad \text{for all $t \in \T$.}
$$

Suppose by contradiction that $|\T| \ge M + 1$. Since the requests are in
general position, the augmented matrix of this system has rank $M+1$. By
Lemma~\ref{lemma:overdetermined-system-of-linear-equations} with $L = |\T|$,
this system of equations has no solution. This contradicts the assumption that
$\lambda_1, \lambda_2, \dots, \lambda_M$ is a solution of this system of
equations.
\end{proof}

\begin{lemma}[Lagrangian policies are close to optimal]
\label{lemma:lagrangian-policies-are-close-to-optimal}
Suppose that the requests $R_1, R_2, \dots, R_T$ are in general position. Let
$\mu_1, \dots, \mu_T$, $\lambda_1, \dots, \lambda_M$ be an optimal solution of
the dual program. Let $x_{1,1}, \dots, x_{T,N}$ be any optimal solution of the
primal program. Let $\lambda = (1, - \lambda_1, - \lambda_2, \dots, -
\lambda_M)$ and let $\pi_\lambda$ be the corresponding Lagrangian policy. Then,
for every $j=0,1,\dots,M$,
$$
\frac{1}{T} \left| \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_\lambda(R_t)_i - x_{t,i} \right) \right| \le \frac{2M}{T} \: .
$$
\end{lemma}

\begin{proof}
Let $\T$ be the set of indices $t$ for which the argmax is not unique. Formally,
$$
\T = \left\{ t \in \Z ~:~ 1 \le t \le T, \text{argmax \eqref{equation:argmax} is \emph{not} unique} \right\}
$$
Theorem~\ref{theorem:few-ties} implies that $|\T| \le M$.

Let $\T^c = \left\{ t \in \Z ~:~ 1 \le t \le T, t \not \in \T \right\}$ be the complement of $\T$.
Proposition~\ref{proposition:computing_primal_from_dual} implies that $x_{t,i} =
\pi(R_t)_{i}$ for all $t \in \T^c$ and all $i = 1,2,\dots,N$.

Thus,
\begin{align*}
& \frac{1}{T} \left| \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_\lambda(R_t) - x_{t,i} \right) \right| \\
& \qquad = \frac{1}{T} \left| \sum_{t \in \T}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_\lambda(R_t)_i - x_{t,i}  \right) + \sum_{t \in \T^c}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_\lambda(R_t)_i - x_{t,i} \right) \right|  \\
& \qquad = \frac{1}{T} \left| \sum_{t \in \T}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_\lambda(R_t)_i - x_{t,i} \right) \right| \\
& \qquad \le \frac{1}{T} \sum_{t \in \T}^T \sum_{i=1}^N \left| R_{t,i,j} \left( \pi_\lambda(R_t)_i - x_{t,i} \right) \right| \\
& \qquad \le \frac{1}{T} \sum_{t \in \T}^T \left( \max_{i=1,2,\dots,N} \left| R_{t,i,j} \right| \right) \cdot \left( \sum_{i=1}^N \left| \left( \pi_\lambda(R_t)_i - x_{t,i} \right) \right| \right) \\
& \qquad \le \frac{1}{T} \sum_{t \in \T}^T \left( \max_{i=1,2,\dots,N} \left| R_{t,i,j} \right| \right) \cdot \left( \sum_{i=1}^N \left| \pi_\lambda(R_t)_i \right| + \left| x_{t,i} \right| \right) \\
& \qquad \le \frac{2|\T|}{T} \\
& \qquad \le \frac{2M}{T} \: .
\end{align*}

\end{proof}

\section{Optimization problem for the expected values}

Recall the optimization problem over all possible measurable policies $\pi:\Requests \to \Delta_N$.

\begin{equation}
\label{equation:distribution-policy-optimization-2}
\begin{aligned}
\maximize_{\pi \in \Pi} \qquad & \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,0} \right] \\
\subjectto \qquad & \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,j} \right] \le B_j & \text{for all $j=1,2,\dots,M$} \\
\end{aligned}
\end{equation}

The problem might be infeasible. However, if it feasible and and the
distribution $D$ has a probability density (with respect to Lebesgue measure on
$\Requests$), then there exists an optimal solution that is a Lagrangian policy.

\begin{lemma}[Concentration inequality]
\label{lemma:concentration-inequality}
Let $D$ be any distribution over $\Requests$. Let $\delta \in (0,1)$. Let $T$ be
a positive integer. Let $j$ be an integer such that $0 \le j \le M$.
Let $\pi: \Requests \to \Delta_N$ be an arbitrary measurable policy.
Suppose $R, R_1, R_2, \dots, R_T$ is an i.i.d. sample from $D$.
Then, with probability at least $1-\delta$,
$$
\left| \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,j} \right]  -  \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi(R_t)_i  \right|
\le \sqrt{ \frac{2\log(2/\delta)}{T}}
$$
\end{lemma}

\begin{proof}
Let $X_t = \sum_{i=1}^N \pi(R_t)_i \cdot R_{t,i,j}$ for all $t=1,2,\dots,T$. Let
$Y = \frac{1}{T} \sum_{t=1}^T X_t$. Since $X_t$ lies in the interval $[-1,1]$,
by Hoeffding's inequality,
$$
\Pr \left[ \left| \Exp[Y] - Y \right| \ge \epsilon \right]  \le 2 \exp\left( - \frac{\epsilon^2 T}{2} \right) \: .
$$
We set $\delta = 2 \exp\left( - \frac{\epsilon^2 T}{2} \right)$ and
solve for $\epsilon$. We get that with probability at least $1-\delta$,
$$
\left| \Exp[Y] - Y \right| \le \sqrt{ \frac{2\log(2/\delta)}{T}} \: .
$$
The theorem follows from the fact that
$$
\Exp[Y] = \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,j}  \right]
\quad \text{and} \quad
Y = \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi(R_t)_i \: .
$$
\end{proof}

\begin{theorem}[Lagrangian policies are optimal]
Let $D$ be a probability distribution over $\Requests$ with a probability
density. Suppose that the optimization problem has a feasible solution, i.e.,
there exists a measurable policy $\pi: \Requests \to \Delta_N$ that satisfies
all the constraints in~\eqref{equation:distribution-policy-optimization-2}.
Then, there exists $\lambda \in \R^{M+1}$ such that $\|\lambda\|_2 = 1$ and the
Lagrangian policy $\pi_\lambda$ is an optimal solution of the optimization
problem~\eqref{equation:distribution-policy-optimization-2}.
\end{theorem}

\begin{proof}
Let $V$ be the optimal value of the optimization problem. That is,
$$
V = \sup \left\{ \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,0} \right] ~:~ \pi \in \Pi, \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi(R)_i \cdot R_{i,j} \right] \le B_j \text{ for all $j=1,2,\dots,M$} \right\} \: .
$$
Since the optimization problem is feasible, $V > -\infty$ and, in particular, $V \in [-1,1]$.

Fix $\epsilon > 0$. Since the optimization problem is feasible, there exists a
measurable policy $\pi^*: \Requests \to \Delta_N$ that is feasible and
$\epsilon$-optimal. That is,
\begin{equation}
\label{inequality:optimal-policy-feasibility}
\Exp_{R \sim D} \left[ \sum_{i=1}^N \pi^*(R)_i \cdot R_{i,j} \right] \le B_j \qquad \text{for all $j=1,2,\dots,M$} \: .
\end{equation}
and
\begin{equation}
\label{inequality:optimal-policy-optimality}
\Exp_{R \sim D} \left[ \sum_{i=1}^N \pi^*(R)_i \cdot R_{i,0} \right] \ge V - \epsilon \: .
\end{equation}

Let $R_1, R_2, \dots$ be an infinite sequence of requests sampled i.i.d. from
$D$. Let $T$ be a positive integer. Let
\begin{align*}
\delta_T & = \frac{1}{(M+1)\pi^2 T^2} \\
B_{T,j}' & = B_j + \sqrt{ \frac{2\log(2/\delta_T)}{T}} \qquad \text{for all $j=1,2,\dots,M$.}
\end{align*}
We consider the modified versions of the primal and dual linear
programs~\eqref{equation:primal-linear-program}
and~\eqref{equation:dual-linear-program} with $B_{T,j}'$ instead of $B_j$,
\begin{equation}
\label{equation:primal-linear-program-modified}
\begin{aligned}
\maximize_{x_{1,1}, \dots, x_{T,N} \in \R} \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot x_{t,i} \\
\subjectto \qquad & \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot x_{t,i} \le B_{T,j}' && \qquad \text{for all $j=1,2,\dots,M$} \\
& \sum_{i=1}^N x_{t,i} = 1 && \qquad \text{for all $t=1,2,\dots,T$} \\
& x_{t,i} \ge 0 && \qquad \text{for all $i=1,2,\dots,N$} \\
& && \qquad \text{and all $t=1,2,\dots,T$}
\end{aligned}
\end{equation}
and
\begin{equation}
\label{equation:dual-linear-program-modified}
\begin{aligned}
\minimize_{\substack{\mu_1, \dots, \mu_T \in \R \\ \lambda_1, \dots, \lambda_M \in \R}} \qquad & \sum_{t=1}^T \mu_t + \sum_{j=1}^M B_{T,j}' \lambda_j \\
\subjectto \qquad
& \mu_t + \frac{1}{T} \sum_{j=1}^M R_{t,i,j} \lambda_j  \ge \frac{1}{T} R_{t,i,0} && \qquad \text{for all $i=1,2,\dots,N$} \\
& && \qquad \text{and all $t=1,2,\dots,T$} \\
& \lambda_j \ge 0 && \qquad \text{for all $j=1,2,\dots,M$}
\end{aligned}
\end{equation}
Finally, let $\delta_T' = \frac{1}{\pi^2 T^2}$.

Let $\E_1$ be the probability event that for all infinitely many values of $T=1,2,\dots$
and for all $j=0,1,2,\dots,M$,
\begin{equation}
\label{inequality:optimal-policy-concentration}
\left| \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi^*(R)_i \cdot R_{i,j} \right] - \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi^*(R_t)_i  \right|
\le \sqrt{ \frac{2\log(2/\delta_T)}{T}} \: .
\end{equation}
Let $\E_2$ be the probability event that for all infinitely many values of $T=1,2,\dots$,
for all $j=0,1,2,\dots,M$, and for all $\lambda \in \R^{M+1}$,
\begin{multline}
\label{inequality:uniform-convergence}
\left| \Exp_{R \sim D} \left[ \sum_{i=1}^N \pi_\lambda(R)_i \cdot R_{i,j}  \right]  -  \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi_\lambda(R_t)_i  \right|
\\
\le \sqrt{\frac{8(M+1) (\ln(4T) + 2 \ln N)}{T}} + \sqrt{ \frac{2\log(1/\delta'_T)}{T} } \: .
\end{multline}
Let $\E_3$ the probability event that, for all infinitely many values of
$T=1,2,\dots$, the requests $R_1, R_2,\dots,R_T$ are in general position. Let
$\E$ be the intersection of the events $\E_1$, $\E_2$, $\E_3$.

For a fixed $T$ and a fixed $j$, by Lemma~\ref{lemma:concentration-inequality},
the inequality \eqref{inequality:optimal-policy-concentration} holds with
probability at least $1 - \delta_T$. Since $\sum_{j=0}^M \sum_{T=1}^\infty
\delta_T = \sum_{j=0}^M \sum_{T=1}^\infty \frac{1}{(M+1)\pi^2 T^2} =
\frac{1}{6}$, by union bound, the probability of event $\E_1$ is at least $5/6$.
For a fixed $T$ and all $j=0,1,2,\dots,M$ and $\lambda \in \R^{M+1}$, by
Theorem~\ref{theorem:uniform-convergence}, the inequality
\eqref{inequality:uniform-convergence} with probability at least $1 -
\delta_T'$. Since Since $\sum_{T=1}^\infty \delta_T' = \sum_{T=1}^\infty
\frac{1}{\pi^2 T^2} = \frac{1}{6}$, by union bound, the probability of event
$\E_2$ is at least $5/6$. For a fixed $T$, the requests $R_1, R_2,\dots,R_T$ are
in general position with probability one. By union bounds, the probability of
event $\E_3$ is also one. Finally, the probability of event $\E$ is at least $1
- 1/6 - 1/6 - 0 = 2/3$.

For the rest proof, we fix $R_1, R_2, \dots$ to be a particular non-random
sequence that belongs to the probability event $\E$.

We combine inequalities~\eqref{inequality:optimal-policy-concentration},
\eqref{inequality:optimal-policy-feasibility} and the definition of $B_{T,j}'$
and we get
\begin{equation}
\label{inequality:optimal-policy-feasibility-on-sample}
\frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot \pi^*(R_t)_i \le B_{T,j}'
\qquad \text{for all $j=1,2,\dots,M$ and all $T=1,2,\dots$.}
\end{equation}
Similarly, we combine inequalities~\eqref{inequality:optimal-policy-concentration},
and \eqref{inequality:optimal-policy-optimality} and get that
\begin{equation}
\label{inequality:optimal-policy-optimality-on-sample}
\frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot \pi^*(R_t)_i \ge V - \epsilon - \sqrt{ \frac{2\log(2/\delta_T)}{T}}
\qquad \text{for all $T=1,2,\dots$.}
\end{equation}

The inequality~\eqref{inequality:optimal-policy-feasibility-on-sample} implies
that $x_{t,i} = \pi^*(R_t)_i$, $i=1,2,\dots,N$, $t=1,2,\dots$ is a feasible
solution of the modified primal linear
program~\eqref{equation:primal-linear-program-modified}. In particular, this
means that the modified primal linear
program~\eqref{equation:primal-linear-program-modified} is feasible.

Let $x^{(T)}_{t,i}$, $t=1,2,\dots,T$, $i=1,2,\dots,N$, be an optimal solution of
the modified primal program~\eqref{equation:primal-linear-program-modified}.
Since $x_{t,i} = \pi^*(R_t)_i$ is a feasible solution of the modified primal
program~\eqref{equation:primal-linear-program-modified},
\begin{equation}
\label{inequality:sample-optimal-policy-beats-optimality-on-sample}
\frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot x^{(T)}_{t,i}
\ge \frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot \pi^*(R_t)_i
\qquad \text{for all $T=1,2,\dots$.}
\end{equation}
We combine the last inequality with inequality~\eqref{inequality:optimal-policy-optimality-on-sample}
and get that
\begin{equation}
\label{inequality:sample-optimal-policy-is-close-optimal}
\frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,0} \cdot x^{(T)}_{t,i}
\ge V - \epsilon - \sqrt{ \frac{2\log(2/\delta_T)}{T}}
\qquad \text{for all $T=1,2,\dots$.}
\end{equation}
The feasibility of the $x^{(T)}_{t,i}$, $t=1,2,\dots,T$, $i=1,2,\dots,N$ implies that
\begin{equation}
\label{inequality:sample-optimal-policy-feasibility}
\frac{1}{T} \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \cdot x^{(T)}_{t,i} \le B_{T,j}'
\qquad \text{for all $j=1,2,\dots,M$ and all $T=1,2,\dots$.}
\end{equation}

Since the modified primal
program~\eqref{equation:primal-linear-program-modified} has a bounded feasible
set and it is feasible, its dual~\eqref{equation:dual-linear-program-modified}
is feasible and has an optimal solution. Let $\lambda_1^{(T)}, \dots,
\lambda_M^{(T)}$, $\mu_1^{(T)}, \dots, \mu_T^{(T)}$ be an optimal solution of
the dual linear program~\eqref{equation:dual-linear-program-modified}. Let
$\lambda^{(T)} = (1, -\lambda_1^{(T)}, -\lambda_2^{(T)}, \dots,
-\lambda_M^{(T)})$ and let $\pi_{\lambda^{(T)}}$ be the corresponding Lagrangian
policy. Since the requests $R_1, R_2, \dots, R_T$ are in general position,
Lemma~\ref{lemma:lagrangian-policies-are-close-to-optimal} implies that
\begin{equation}
\label{inequality:lagrangian-policy-rounding-error}
\frac{1}{T} \left| \sum_{t=1}^T \sum_{i=1}^N R_{t,i,j} \left( \pi_{\lambda^{(T)}}(R_t)_i - x^{(T)}_{t,i} \right) \right| \le \frac{2M}{T} \: .
\end{equation}
We combine inequality~\eqref{inequality:lagrangian-policy-rounding-error} with
inequalities~\eqref{inequality:sample-optimal-policy-is-close-optimal}
and \eqref{inequality:uniform-convergence}. We get that
\begin{multline}
\label{inequality:lagrangian-policy-optimality}
\Exp_{R \sim D} \left[ \sum_{i=1}^N \pi_{\lambda^{(T)}}(R_t)_i \cdot R_{i,j}  \right] \ge V - \epsilon - \sqrt{ \frac{2\log(2/\delta_T)}{T}} - \sqrt{\frac{8(M+1) (\ln(4T) + 2 \ln N)}{T}} \\
- \sqrt{ \frac{2\log(1/\delta'_T)}{T} } - \frac{2M}{T}  \: .
\end{multline}
Similarly, we combine the inequality~\eqref{inequality:lagrangian-policy-rounding-error}
with inequalities~\eqref{inequality:sample-optimal-policy-feasibility}
and \eqref{inequality:uniform-convergence} and definition of $B_{T,j}'$. We get that
\begin{multline}
\label{inequality:lagrangian-policy-feasibility}
\Exp_{R \sim D} \left[ \sum_{i=1}^N \pi_{\lambda^{(T)}}(R_t)_i \cdot R_{i,j}  \right]
\le B_j + \sqrt{ \frac{2\log(2/\delta_T)}{T}} + \frac{2M}{T}
+ \sqrt{\frac{8(M+1) (\ln(4T) + 2 \ln N)}{T}} \\
+ \sqrt{ \frac{2\log(1/\delta'_T)}{T} }
\: .
\end{multline}

Since $\lambda^{(T)}$ is a non-zero vector, we can normalize it to unit length
and the inequalities~\eqref{inequality:lagrangian-policy-optimality}
and~\eqref{inequality:lagrangian-policy-feasibility} continue to hold.
Thus, after normalization, $\lambda^{(1)}, \lambda^{(2)}, \dots$
is a sequence of unit length vectors in $\R^{M+1}$. Since the unit sphere
is a compact set, there exists

\end{proof}

\appendix

\section{Technical lemmas}

\begin{lemma}[Maximum of sub-Gaussian variables]
\label{lemma:maximum-of-sub-gaussian-variables}
Let $\sigma \ge 0$ and let $K \ge 1$. Let $Y_1, Y_2, \dots, Y_K$ be random
variables that satisfy the $\sigma$-sub-Gaussian property
$$
\Exp\left[ e^{sY_k} \right] \le e^{s^2\sigma^2/2} \qquad \text{for all $s \in \R$ and for all $k=1,2,\dots,K$.}
$$
Then,
$$
\Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le \sigma \sqrt{2 \ln(2K)} \: .
$$
\end{lemma}

\begin{proof}
For any $s \in \R$,
\begin{align*}
\exp\left(s \Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \right)
& = \exp\left(s \Exp\left[ \max_{k=1,2,\dots,K} \max \left\{Y_k, -Y_k\right\} \right] \right) \\
& \le \Exp\left[ \exp\left(s \max_{k=1,2,\dots,K} \max \left\{Y_k, -Y_k\right\} \right) \right] \\
& = \Exp\left[ \max_{k=1,2,\dots,K} \max \left\{e^{sY_k}, e^{-sY_k}\right\} \right] \\
& \le \Exp\left[ \sum_{k=1}^K \left( e^{sY_k} + e^{-sY_k} \right) \right] \\
& = \sum_{k=1}^K \left( \Exp\left[ e^{sY_k} \right] + \Exp\left[ e^{-sY_k} \right] \right) \\
& \le \sum_{k=1}^K \left( e^{s^2\sigma^2/2} + e^{s^2\sigma^2/2} \right) \\
& = 2K e^{s^2\sigma^2/2} \: .
\end{align*}
The first equality follows from that $|Y_k| = \max \left\{Y_k, -Y_k\right\}$.
The first inequality follows from Jensen's inequality for the convex function $x
\mapsto e^{sx}$. The second inequality follows from the fact that maximum of
non-negative numbers is upper bounded by their sum. The last inequality follows from
the sub-Gaussian property for $Y_k$.

Thus, for any $s \in \R$,
$$
\exp\left(s \Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \right) \le 2K e^{s^2\sigma^2/2} \: .
$$
Taking logarithms, we get that for any $s \in \R$,
$$
s \Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le \frac{s^2 \sigma^2}{2} + \ln(2K) \: .
$$
We restrict ourselves to $s > 0$ and divide both sides by $s$. We get that for any $s > 0$,
$$
\Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le \frac{s \sigma^2}{2} + \frac{\ln(2K)}{s} \: .
$$
If $\sigma > 0$, we choose $s = \frac{2\sqrt{\ln(2K)}}{\sigma}$ and get that
$$
\Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le \frac{s \sigma^2}{2} + \frac{\ln(2K)}{s} = \sigma \sqrt{2 \ln(2K)} \: .
$$
If $\sigma = 0$ then for any $s > 0$,
$$
\Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le \frac{\ln(2K)}{s} \: .
$$
Taking $s \to \infty$, we get that $\Exp\left[ \max_{k=1,2,\dots,K} |Y_k| \right] \le 0$.
\end{proof}

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}
